On the discovery and origin of $r$-process stars in the Milky Way

### 2018-02-07 article/ms.tex
edits up to early discussion; rest in person

	\begin{abstract}
		The production of heavy elements occurs through the slow and rapid neutron capture process. Numerous sites satisfy the conditions needed for the r-process to occur, but the relative frequency of these sites is unknown largely due to the lack of a large number of stars with strong r-process signatures. Here we report the discovery of 47 r-process enhanced stars discovered in the LAMOST survey, more than doubling the known number in our Galaxy. To our knowledge, we provide the first estimate on whether neutron star mergers are frequent enough to account for all known r-process stars in the Milky Way. \todo{We conclude....}. \todo{The discovery of these stars implies multiple r-process sites.}
		% contrary to previous conclusions in the literature including all r-process material being produced in small amounts in common events such as core-collapse supernovae, or the material exclusively being produced in very large amounts in very rare events, such as neutron star mergers. 
	\end{abstract}

@@

	Heavy elements ($Z > 30$) are synthesised through the slow (s) and rapid (r) neutron capture processes \citep{Sneden2008}. The nucleosynthesis of these elements has been a key area in astrophysics research since it was first theorised  \citep{Burbidge1957}.

	Potential sites of the r-process have been discussed for over sixty years. Rapid neutron capture occurs when the neutron density is high enough that the rate of neutron capture will occur faster than the rate at which the isotope undergoes a $\beta$-decay. The rapid neutron capture sites are very specific, with the most favoured being core-collapse supernovae, neutron star mergers, or magnetorotationally driven supernovae. Core-collapse supernovae are relatively common throughout the Galaxy, but they only produce small amounts of r-process material. Neutron star mergers or magnetorotationally driven supernovae, however, occur much less frequently than core-collapse supernovae, but produce far more r-process material. Events such as these are theorised to account for the case of the highly r-process enhanced ultra-faint dwarf (UFD) galaxy Reticulum II \citep{Ji2016}. %Neutron star mergers may have a stronger argument to be the parent site of this level of r-process enhancement in such a small volume. However, magnetorotationally driven supernovae are also a possible solution.
Magnetorotationally driven supernovae may not produce as much r-process material as neutron star mergers (although still more than core-collapse supernovae), they are thought to occur more frequently than neutron-star mergers \todo{[citation needed]}, and the predicted abundance patterns make them observationally indistinguishable from neutron star mergers \citep{Ji2016}.

	Europium ($Z=63$), is entirely produced through the r-process. This makes it a very useful signature for identifying stars that show evidence of excess pollution by r-process material. %\todo{CLUMSY, lose the first half-sentence:The recent discovery of the UFD Reticulum II used Europium as evidence of its r-process enhancement, and found the particularly interesting result that the UFD's level of r-process enrichment was too high to have occurred through a core-collapse supernovae event.}
	Historically, UFD galaxies have demonstrated some of the lowest abundance ratios of r-process elements seen in the Milky Way or its satellite systems. However, the analysis of several of Reticulum II's stars show they are greatly enriched, 2-3 orders of magnitude higher than stars found in any other UFD galaxy \citep{Ji2016}. This degree of enrichment supports the theory that a single rare event must have occurred early on. This is supported by the observed heavy element yields which are found to be some 1000 times higher than what is achievable on average from core-collapse supernovae ejecta. It is highly improbable that 1000 supernovae could have contributed to Reticulum II, suggesting that the r-process material must be produced by a rarer astrophysical event such as a binary neutron star merger.

	While neutron star mergers or magnetorotationally driven supernovae are likely to produce significant amounts of r-process material, it does not suggest that all r-process material stems from these sources either. Rather, the rarity of these events possibly indicates the rarity of stars found to be heavily enhanced in these elements. One could ask if neutron-star mergers could be responsible for all r-process stars known in the Milky Way? On the other hand, if we assume very little gas mixing, could could core-collapse supernovae create enough material to explain a single r-process star? If neither of these prove to be true then we can confirm that there must be multiple sties to the r-process, providing a step forward in understanding the relative fraction of these mechanisms in the Milky Way.

%	 stars we've found? And if neither of these prove to be the case, can we then conclude that there must be multiple sites of r-process element creation throughout the Milky Way, that create these elements to different extents?

	To answer these questions it is necessary to identify a large sample of r-process stars.
	 However, r-process stars are very rare, which makes finding this stars difficult. For this reason we must look to massive data sets, such as the LAMOST sky survey. In this letter we report the  discovery of 47 r-process enhanced stars identified from LAMOST.
	In Section 2 we outline the methods used to identify r-process enhanced candidates, and our analysis methods. In Section 3 we discuss what our results imply about the sites of r-process. We provide concluding remarks in Section 4.

@@

	We identified candidate r-process enhanced stars by searching for deviations surrounding between the pseudo-continuum-normalised flux and the best-fitting model spectrum surrounding

	 flux and the pseudo-continuum-normalised spectra

	 with an over-abundance of Europium by searching for deviations from the normalized spectra supplied by \textit{The Cannon}. A negative deviation implies an enhancement in a particular element while a positive deviation implies a depletion (compared to what the data-driven model predicted).

	At two absorption lines for Europium (4129\,\AA\ and 4205\,\AA) we fitted Gaussian profiles to each star in the 454,180 giants from LAMOST DR2, and we recorded the amplitude and the wavelength of each, as well as their corresponding errors. Candidate Europium enhanced stars were selected by identifying stars that met the following criteria:

	\begin{itemize}
		\item The amplitude of their deviation from the model needed to be less than $\rm{A}<-0.05$;
		\item The signal to noise (S/N) must be greater than 30;
		\item The deviation from the required wavelength in question must be less than 2 \AA;
		\item The $\chi_{\rm{r}}^{2}$ value (where the $\chi_{\rm{r}}^{2}$ value represents how well the model fit the spectrum) must be less than 3, $\chi_{\rm{r}}^{2}<3$.
		\item All of these conditions must be met for both 4129 \AA\ and 4205 \AA\ for us to consider them sufficiently Eu enhanced.
	\end{itemize}

	We gathered an initial sample of 62 candidates out of the 454180 LAMOST giant sample. These were then visually scrutinised to identify those that lacked visual evidence that they were enhanced. After this process we were left with 47 potentially r-process enhanced candidates, 22 of which stand out as very strong candidates.

	\section{The r-process}
	For the r-process to occur there must be a very high neutron density, such that neutron capture must be able to occur before the nucleus undergoes a $\beta$-decay (assuming the nucleus is in an unstable state). The conditions that can create this are rare. Events such as supernovae or neutron star mergers list among the most favoured sites.

	\subsection{Selection effects}
	The LAMOST giants we analyse belong to a broad range of metallicity ($\rm{[Fe/H]} -1.6\ \rm{to}\ 0.6$), and don't cluster towards any position in the observed sky. These characteristics argue that the abundance we see could not have originated from a common era or site.

	The previous sample of r-process stars, before the release of this letter, almost exclusively consists of metal-poor stars. This fact is not to be confused with saying that most r-process stars found should be metal-poor stars, rather that at low metallicities we know that the s-process will not contribute, so neutron capture elements formed from these stars would be only from the r-process. Over half the current known r-process stars come from a very metal poor ultra faint dwarf galaxy, Reticulum II. However, Europium we know to be totally made from the r-process, therefore even at metallicities where the s-process contributes, Europium is a valuable indicator of r-process synthesis.

	\subsection{Multiple r-process sites?}
	While it is possible that all currently known r-process enhancements can be made by neutron star mergers, we still observe stellar populations in the galaxy which show particularly weak enrichment from r-process elements ([Eu/Fe] < 0.7). These enhancements imply that an additional site of r-process element synthesis is likely to be active. As we show in \ref{rates}  neutron star merger rates may show that all heavily enriched stars can be enhanced by these phenomena within an adequate time-scale. However, we are not able to assume that as a consequence of the merger some stars are greatly enhanced, while others are only very slightly enhanced. It is more accurate to propose that at least two sites exist, with one (potentially mergers) producing a massive amount of enhanced material, and another only producing a small amount, such as supernovae.

@@

	Regardless of the exact mechanism, core-collapse supernovae seem to release $M_{\rm{Eu}}\approx10^{-7.5} M_{\odot}$ per event \citep{Argast2004} . While this may not be a particularly significant amount, core-collapse supernovae occur frequently $44700\ \rm{Gpc}^{-3} \rm{yr^{-1}}$ \citep{Li2011}, and so it isn't unreasonable to suggest that, with very little mixing occurring, that much if not all r-process material could have been formed by these and recycled in the stars we present in this letter.

>>>>

	\begin{abstract}
		The production of heavy elements occurs through the slow and rapid neutron capture process. Numerous sites satisfy the conditions needed for the r-process but the relative frequency of these sites is unknown, largely due to the lack of a large number of stars with strong r-process signatures. Here we report the discovery of 47 r-process enhanced stars discovered in the LAMOST survey, more than doubling the known number in our Galaxy. To our knowledge, we provide the first estimate on whether neutron star mergers are frequent enough to account for all known r-process stars in the Milky Way. \todo{We conclude....}. \todo{The discovery of these stars implies multiple r-process sites.}
		% contrary to previous conclusions in the literature including all r-process material being produced in small amounts in common events such as core-collapse supernovae, or the material exclusively being produced in very large amounts in very rare events, such as neutron star mergers. 
	\end{abstract}

@@

	Heavy elements ($Z > 30$) are synthesised through the slow (s) and rapid (r) neutron capture processes \citep[e.g.,][]{Burbidge1957,Sneden2008}. 
	%Potential sites of the r-process have been discussed for over sixty years. 
	Rapid neutron capture occurs when the neutron density is high enough that the rate of neutron capture will occur faster than the rate at which the isotope undergoes a $\beta$-decay. Extremely high neutron densities (\todo{of how much}) are necessary for the r-process to occur. Sites meeting this criteria include core-collapse supernovae, neutron star mergers, or magnetorotationally driven supernovae. Core-collapse supernovae are relatively common throughout the Galaxy, but they only produce small amounts of r-process material. Neutron star mergers or magnetorotationally driven supernovae, however, occur much less frequently than core-collapse supernovae, but produce far more r-process material. Events such as these are theorised to account for the case of the highly r-process enhanced stars in the ultra-faint dwarf (UFD) galaxy Reticulum II \citep{Ji2016}. %Neutron star mergers may have a stronger argument to be the parent site of this level of r-process enhancement in such a small volume. However, magnetorotationally driven supernovae are also a possible solution.
Magnetorotationally driven supernovae may not produce as much r-process material as neutron star mergers (although still more than core-collapse supernovae), they are thought to occur more frequently than neutron-star mergers \todo{[citation needed]}, and the predicted abundance patterns make them observationally indistinguishable from neutron star mergers \citep{Ji2016}.

	Europium ($Z=63$) is entirely produced through the r-process. This makes it a very useful signature for identifying stars that show evidence of excess pollution by r-process material. %\todo{CLUMSY, lose the first half-sentence:The recent discovery of the UFD Reticulum II used Europium as evidence of its r-process enhancement, and found the particularly interesting result that the UFD's level of r-process enrichment was too high to have occurred through a core-collapse supernovae event.}
	Historically, UFD galaxies have demonstrated some of the lowest abundance ratios of r-process elements seen in the Milky Way or its satellite systems. However, the analysis of several of Reticulum II's stars show they are greatly enriched, 2-3 orders of magnitude higher than stars found in any other UFD galaxy \citep{Ji2016}. This degree of enrichment supports the theory that a single rare event must have occurred early on in the formation history of Reticulum II. This is supported by the observed heavy element yields, which are found to be some 1000 times higher than what is achievable on average from core-collapse supernovae ejecta. It is highly improbable that 1000 supernovae could have contributed to Reticulum II, suggesting that the r-process material must be produced by a rarer astrophysical event such as a binary neutron star merger.

	While neutron star mergers or magnetorotationally driven supernovae are likely to produce significant amounts of r-process material, it does not suggest that all r-process material stems from these sources either. Rather, the rarity of these events might imply that they are also responsible for the rare numbers of stars  observed to have significant enhancements in r-process material. One could ask if neutron-star mergers could be responsible for all r-process stars known in the Milky Way? On the other hand, if we assume very little gas mixing, could could core-collapse supernovae create enough material to explain a single r-process star? If neither of these prove to be true then we can confirm that there must be multiple sties to the r-process, providing a step forward in understanding the relative fraction of these mechanisms in the Milky Way.

%	 stars we've found? And if neither of these prove to be the case, can we then conclude that there must be multiple sites of r-process element creation throughout the Milky Way, that create these elements to different extents?
	In order to answer these questions it is necessary to identify a large sample of r-process stars.
	 However, r-process stars are very rare, which makes finding this stars difficult. For this reason we must look to massive data sets, such as the LAMOST sky survey. In this letter we report the discovery of 47 r-process enhanced stars identified from LAMOST.
	In Section 2 we outline the methods used to identify r-process enhanced candidates, and our analysis methods. In Section 3 we discuss what our results imply about the sites of r-process. We provide concluding remarks in Section 4.

@@

	We identified candidate r-process enhanced stars by searching for deviations surrounding between the pseudo-continuum-normalised flux and the best-fitting data-driven model spectrum. Specifically, we looked for deviations at the 4129\,\AA\ and 4205\,\AA\ europium absorption lines. At each transition we fitted a Gaussian profile to the flux residuals for all 454,180 giant stars. We recorded the amplitude, wavelength, and width of each profile, as well as the corresponding uncertainties on these quantities. 
	 Candidate r-process stars were identified as stars that met the following criteria:

	\begin{itemize}
		\item The amplitude of the gaussian profile must be less than $\rm{A}<-0.05$;
		\item The signal to noise (S/N) must be greater than 30;
		\item The deviation from the required wavelength in question must be less than 2 \AA;
		\item The  $\chi_{\rm{r}}^{2}$ value must be less than 3, $\chi_{\rm{r}}^{2}<3$.
		\item All of these conditions must be met for both 4129 \AA\ and 4205 \AA\ for us to consider them sufficiently Eu enhanced.
	\end{itemize}

	These filters produced an initial sample of 62 candidates from the total sample of 454,180. An example candidate is shown in Figure \ref{fig:starindex_423451}. We visually scrutinised each candidate to remove false positives. After this process we were left with 47 likely r-process enhanced candidates, 22 of which stand out as very strong candidates.

	\section{The r-process}
	% Repeating the introduction:
	%For the r-process to occur there must be a very high neutron density, such that neutron capture must be able to occur before the nucleus undergoes a $\beta$-decay (assuming the nucleus is in an unstable state). The conditions that can create this are rare. Events such as supernovae or neutron star mergers list among the most favoured sites.

	The LAMOST giants we analyse belong to a broad range of metallicity ($-1.6 < \rm{[Fe/H]} < 0.6$), and don't cluster in any part of the sky, suggesting they do not belong to a single stellar system.

	The previous sample of r-process stars, before the release of this letter, almost exclusively consists of metal-poor stars \todo{[OMG CITATIONS NEEDED DUDE]}. This fact is not to be confused with saying that most r-process stars found should be metal-poor stars, rather that at low metallicities we know that the s-process will not contribute, so neutron capture elements formed from these stars would be only from the r-process. Over half the current known r-process stars are members of the ultra faint dwarf galaxy, Reticulum II. However, Europium we know to be totally made from the r-process, therefore even at metallicities where the s-process contributes, Europium is a valuable indicator of r-process synthesis.

	\subsection{Multiple r-process sites?}
	While it is possible that all currently known r-process enhancements can be made by neutron star mergers, we still observe stellar populations in the galaxy which show particularly weak enrichment from r-process elements ([Eu/Fe] < 0.7). These enhancements imply that an additional site of r-process element synthesis is likely to be active \todo{or there is very efficient mixing?? I think here you just want to say that either there is $>1$ r-process site, or there is extremely heterogeneous levels of mixing from neutron-star mergers in order to produce stars with [Eu/Fe] abundances that range many orders of magnitude...}.

	%As we show in \ref{rates}  neutron star merger rates may show that all heavily enriched stars can be enhanced by these phenomena within an adequate time-scale. However, we are not able to assume that as a consequence of the merger some stars are greatly enhanced, while others are only very slightly enhanced. It is more accurate to propose that at least two sites exist, with one (potentially mergers) producing a massive amount of enhanced material, and another only producing a small amount, such as supernovae.

@@

	Regardless of the exact mechanism, core-collapse supernovae seem to release $M_{\rm{Eu}}\approx10^{-7.5} M_{\odot}$ per event \citep{Argast2004} . While this may not be a particularly significant amount, core-collapse supernovae occur frequently \citep[$44700\ \rm{Gpc}^{-3} \rm{yr^{-1}}$;][]{Li2011}, and so it isn't unreasonable to suggest that, with very little mixing occurring, that much if not all r-process material could have been formed by these and recycled in the stars we present in this letter.
